Physcraper: A Python package for continually updated phylogenetic trees using the Open Tree of Life

### 2021-01-07 sources/bmc_template_submission/bmc_article.tex
links to studys and data packages as citations

Taxon names in the input tree must be standardized to OpenTree taxonomy
\cite{rees2017automated} using OpenTree's bulk Taxonomic Name Resolution Service
\href{https://tree.opentreeoflife.org/curator/tnrs/}{TNRS} tool. Users can upload their
own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's
\href{https://github.com/opentreeoflife/phylesystem}{Phylesystem} that also have
alignments available on TreeBASE \cite{piel2009treebase}.

@@

The Basic Local Alignment Search Tool, BLAST \cite{altschul1990basic} is used for DNA
sequence search on a remote or local GenBank database. It is constrained to a
"search taxon", a taxonomic group in the NCBI taxonomy that is automatically
identified using the OpenTree
\href{<https://github.com/OpenTreeOfLife/germinator/wiki/Taxonomy-API-v3#mrca>}{API}
\cite{rees2017automated}, as the most recent common ancestor of ingroup taxa that is
also a named clade in the NCBI taxonomy (Figure 1).

BLAST is implemented with the `blastn` function \cite{camacho2009blast} and the BioPython
\cite{cock2009biopython} BLAST function from
\href{https://biopython.org/DIST/docs/api/Bio.Blast.NCBIWWW-module.html}{NCBIWWW module}
modified to accept an alternative BLAST address.
Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank.
New sequences are excluded for analysis if they 1) are not in the search taxon;
2) have an e-value above the cutoff (default to 0.00001); 3) fall outside a min
and max length threshold, defined as the proportion of the average length without
gaps of all sequences in input alignment (default values of 80% and 120%,
respectively); 4) or if they are either identical to or shorter than an existing
sequence in the input alignment and they represent the same taxon in OpenTree or
NCBI taxonomy.
An arbitrary maximum number of randomly chosen sequences per taxon are allowed
(default to 5).

@@

An online literature review in June 2020 (Google scholar search for "ilex phylogeny")
reveals that there are several published phylogenies showing relationships within
the hollies \cite{cuenoud2000molecular, manen2010history, setoguchi2000intersectional,
selbach2009new}, but only two have data publicly available \cite{gottlieb2005molecular, yao2020phylogeny}.
\cite{gottlieb2005molecular} made original tree and alignment available in
\href{https://treebase.org/treebase-web/search/study/summary.html?id=1091}{TreeBASE study 1091}.
The tree sampling 41 species was also added to
OpenTree's Phylesystem \href{https://tree.opentreeoflife.org/curator/study/edit/pg_2827/?tab=home}{study pg\_2827}
and it has been integrated into OpenTree's synthetic tree
\href{https://devtree.opentreeoflife.org/opentree/opentree12.3@mrcaott68451ott89474/Ilex-theizans--Ilex-dumosa}{here}.

The most recent \textit{Ilex} tree \cite{yao2020phylogeny} is available in
OpenTree's Phylesystem \href{https://tree.opentreeoflife.org/curator/study/view/ot_1984}{study ot\_1984}
and in the DRYAD repository \cite{yao2020dryad}.
With 175 tips, the \cite{yao2020phylogeny} tree is the best sampled phylogeny yet available for
the hollies.

@@

>>>>

Taxon names in the input tree must be standardized to OpenTree taxonomy
\cite{rees2017automated} using OpenTree's bulk Taxonomic Name Resolution Service
(TNRS) tool \cite{TNRStool}. Users can upload their
own tree, or choose from among the 2, 950 standardized trees stored in OpenTree's
Phylesystem \cite{phylesystemGithub, mctavish2015phylesystem} that also have
alignments available on TreeBASE \cite{piel2009treebase}.

@@

The Basic Local Alignment Search Tool, BLAST \cite{altschul1990basic} is used for DNA
sequence search on a remote or local GenBank database. It is constrained to a
"search taxon", a taxonomic group in the NCBI taxonomy that is automatically
identified using the OpenTree's taxonomic Most Recent Common Ancestor (MRCA) API
\cite{mrcaAPI, rees2017automated}, as the MRCA of all ingroup taxa that is
also a named clade in the NCBI taxonomy (Figure 1).

BLAST is performed using the `blastn` algorithm \cite{camacho2009blast}
implemented in BioPython's \cite{cock2009biopython} NCBIWWW module \cite{ncbiwww}
modified to accept an alternative BLAST address.
Each sequence in the alignment is BLASTed once against all DNA sequences in GenBank.
New sequences are excluded for analysis if they 1) are not in the search taxon;
2) have an e-value above the cutoff (default to 0.00001); 3) fall outside a min
and max length threshold, defined as the proportion of the average length without
gaps of all sequences in input alignment (default values of 80% and 120%,
respectively); 4) or if they are either identical to or shorter than an existing
sequence in the input alignment and they represent the same taxon in OpenTree or
NCBI taxonomy.
An arbitrary maximum number of randomly chosen sequences per taxon are allowed
(default to 5).

@@

An online literature review in June 2020 (Google scholar search for "ilex phylogeny")
reveals that there are several published phylogenies showing relationships within
the hollies \cite{cuenoud2000molecular, manen2010history, setoguchi2000intersectional,
selbach2009new}, but only two have data publicly available \cite{gottlieb2005molecular, yao2020phylogeny}.
\cite{gottlieb2005molecular} made original tree and alignment available in TreeBASE
(study 1091 \cite{treebase1091}). The tree sampling 41 species is also available from
OpenTree's Phylesystem (study pg\_2827 \cite{pg2827}),
and has been integrated into OpenTree's synthetic tree \cite{mrcaott68451ott89474}.
The most recent \textit{Ilex} tree \cite{yao2020phylogeny} is available in
OpenTree's Phylesystem (study ot\_1984 \cite{ot1984}),
and in the DRYAD repository \cite{yao2020dryad}.
With 175 tips, the \cite{yao2020phylogeny} tree is the best sampled phylogeny yet available for
the hollies.

@@

\textbf{TNRS:} Taxonomic Name Resolution Service

\textbf{MRCA:} Most Recent Common Ancestor

\textbf{BLAST:} Basic Local Alignment Search Tool

\textbf{NCBI:} USA National Center for Biodiversity Information
